---
title: "fpca logistic regr"
author: "Riccardo Morandi"
date: "2023-06-25"
output: html_document
---
set up:

```{r}
setwd("/Users/macbookpro/Documents/Applied Statistics/Project/Materiale_appstat/Code/Functional/")
load("~/Documents/Applied Statistics/Project/Materiale_appstat/Code/Functional/PDC_TOT.RData")
load("~/Documents/Applied Statistics/Project/Materiale_appstat/Code/Functional/pazienti.RData")
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/data.RData")
library(tidyverse)
library(tidymodels)
library(fda)
library(fields)
```


```{r}
time <- 1:365
data_W=PDC_TOT

basis <- create.bspline.basis(rangeval=range(time),nbasis=80,norder=5)

data_W.fd.1 <- Data2fd(y = data_W,argvals = time,basisobj = basis)
plot.fd(data_W.fd.1[1:10], main="B-splines")
```


```{r}
pca_W.1 <- pca.fd(data_W.fd.1,nharm=5,centerfns=TRUE)

plot(pca_W.1$values[1:35],xlab='j',ylab='Eigenvalues')
plot(cumsum(pca_W.1$values)[1:35]/sum(pca_W.1$values),xlab='j',ylab='Contribution to the total variance',ylim=c(0.8,1))
```


```{r}
#Explained variance
(pca_W.1$values)[1:5]/sum(pca_W.1$values)
```


```{r}
# Command of the library fda that automatically does these plots
par(mfrow=c(1,3))
plot(pca_W.1, nx=100, pointplot=TRUE, harm=c(1,2,3), expand=0, cycle=FALSE)
```


```{r}
# scatter plot of the scores
par(mfrow=c(1,1))
plot(pca_W.1$scores[,1],pca_W.1$scores[,2],xlab="Scores FPC1",ylab="Scores FPC2",lwd=2)


#Coloro in base a la variabile label_out
codici=pazienti$labelOUT
dim(sequenze_PDC_totale)
col=rep('red',12588)
col[which(codici=='DECEDUTO')]='green'
# scatter plot of the scores
par(mfrow=c(1,1))
plot(pca_W.1$scores[,2],pca_W.1$scores[,3],xlab="Scores FPC1",ylab="Scores FPC2",lwd=2, col=col)

```

```{r}
codes <- pazienti$CODREG
retained_patients <- data[data$COD_REG %in% codes,]
retained_patients$fpca1 <- pca_W.1$scores[,1]
retained_patients$fpca2 <- pca_W.1$scores[,2]
retained_patients$fpca3 <- pca_W.1$scores[,3]
```

```{r}
df <- retained_patients[,c(3,6,7,8,14,20,21,22)]
```

stratified k_fold cross-validation on labelout

```{r}
set.seed(123)
df_split <- initial_split(df, strata = labelOUT)
df_train <- training(df_split)
df_test <- testing(df_split)

set.seed(234)
df_folds <- vfold_cv(df_train, strata = labelOUT)

```

```{r}
glm_spec <- logistic_reg(engine = "glm")

rec_paper <- recipe(labelOUT ~ ., data = df_train) %>% 
  step_normalize(all_numeric_predictors()) %>% step_dummy(all_nominal_predictors())

wf_paper <- workflow(rec_paper,glm_spec)
```

```{r}
perf <- metric_set(accuracy,
                   sensitivity,
                   specificity,
                   precision,
                   recall,
                   roc_auc)

doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_basic <- fit_resamples(wf_paper, df_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_basic)
```


```{r}
collect_predictions(rs_basic) %>% group_by(id) %>% roc_curve(labelOUT,.pred_DECEDUTO) %>% autoplot()
```

fitting the model to the train data data

```{r}
fit_train <- fit(wf_paper,df_train)
```

looking at the coefficients:

```{r}
tidy(fit_train)
```

all of the coefficients are significant

looking at the odds ratios:

```{r}
tidy(fit_train,exponentiate=TRUE)
```

we can see that the ODDS ratio relative to the PDC_TOT is 1.94

### evaluation on test data

```{r}
test_fit <- last_fit(wf_paper,df_split,metrics = perf)
test_fit %>% collect_metrics()
```
the performance does not change o the test data, no overfitting.

let's investigate more:

```{r}
df_preds <- fit_train %>% augment(new_data = df_test)
```

```{r}
df_preds %>% roc_curve(labelOUT,.pred_DECEDUTO) %>% autoplot()
```

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

we can see that the odds ration for the PDC terms are significant and are 1.21 and 1.32

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```
 






