---
title: "logistics regression clusters"
output: html_notebook
---

let's start by fitting a logistic regression on our data
starting from the total adherence measured via PDC

```{r}
library(tidyverse)
library(tidymodels)
library(themis)
library(patchwork)
```

## Explore the data

loading the data

```{r}
load("~/Documents/Applied Statistics/Project/Materiale_appstat/Code/Functional/data_with_clusters.RData")
```

```{r}
glimpse(data_with_clusters)
```

we want to build a logistic regression model using sesso, età, MCS, PDC_tot

we need some preprocesing on this covariates

```{r}
df <- data_with_clusters %>% select(c(SESSO,labelOUT,eta_Min,MCS,cluster))
df$cluster <- as.factor(df$cluster)
df$labelOUT <- relevel(df$labelOUT,"TRONCATO")
```

```{r}
ggplot(data_with_clusters, aes(cluster)) + geom_bar(aes(fill = labelOUT),binwidth = 0.8)
```

# logisic regression 

## the basic model

stratified k_fold cross-validation on labelout

```{r}
set.seed(123)
df_split <- initial_split(df, strata = labelOUT)
df_train <- training(df_split)
df_test <- testing(df_split)

set.seed(234)
df_folds <- vfold_cv(df_train, strata = labelOUT)
```

```{r}
glm_spec <- logistic_reg(engine = "glm")

rec_paper <- recipe(labelOUT ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% 
  step_dummy(all_nominal_predictors())

wf_paper <- workflow(rec_paper %>% step_upsample(labelOUT),glm_spec)
```

```{r}
perf <- metric_set(accuracy,
                   sensitivity,
                   specificity,
                   precision,
                   recall,
                   roc_auc)

doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_basic <- fit_resamples(wf_paper, df_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_basic)
```
this is quite good: accucracy of 84% and ROC .77

```{r}
collect_predictions(rs_basic) %>% group_by(id) %>% roc_curve(labelOUT,.pred_DECEDUTO) %>% autoplot()
```

fitting the model to the train data data

```{r}
fit_train <- fit(wf_paper,df_train)
```

looking at the coefficients:

```{r}
tidy(fit_train)
```

all of the coefficients are significant

looking at the odds ratios:

```{r}
tidy(fit_train,exponentiate=TRUE)
```

we can see that the ODDS ratio relative to the PDC_TOT is 1.94

### evaluation on test data

```{r}
test_fit <- last_fit(wf_paper,df_split,metrics = perf)
test_fit %>% collect_metrics()
```
the performance does not change o the test data, no overfitting.

let's investigate more:

```{r}
df_preds <- fit_train %>% augment(new_data = df_test)
```

```{r}
df_preds %>% roc_curve(labelOUT,.pred_DECEDUTO) %>% autoplot()
```

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = FALSE) %>% arrange(estimate) 
```

we can see that the odds ration for the PDC terms are significant and are 1.21 and 1.32

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>% filter(term != "n_drugs") %>% 
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```
 
 plotting the odds ratio
 
```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% 
  filter(term != "(Intercept)") %>% 
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 1, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = 0.7, alpha = 0.8) +
  geom_point(size = 2.5) +
  labs(y = NULL, x = "Odds ratios from logistic regression")
```

```{r}
odds_ratios <- test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% 
  filter(term != "(Intercept)") %>% filter(term != "n_drugs")

odds_ratios$term <- c("Age", "Sex M","MCS 1", "MCS 2","MCS 3","MCS 4",
                      "Cluster 3","Cluster 2","Cluster 4")
```


better plot:
```{r}
P1 <- odds_ratios %>% 
ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 1, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = 0.7, alpha = 0.8) +
  geom_point(size = 2.5) +
  labs(y = NULL, x = NULL) + ggtitle("Odds ratios from logistic regression") + 
  theme(legend.position = "none",axis.text = element_text(size = 12),
        plot.title = element_text(size = 16,hjust = 0.5)) 
```

```{r}
P1
ggsave("odds_ratios_pdc_cluster.png", dpi = 1000)
```

```{r}
odds_ratios$color <- as.factor(c(0,0,0,0,0,0,2,3,4))

mycolors <- c("gray0","#00BFC4" ,"#7CAE00", "#C77CFF")

P1 <- odds_ratios %>% 
ggplot(aes(estimate, fct_reorder(term, estimate),color = color)) +
  geom_vline(xintercept = 1, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = 0.7, alpha = 0.8) +
  geom_point(size = 2.5) +
  scale_color_manual(values=mycolors) +
  labs(y = NULL, x = NULL) + ggtitle("Odds ratios from logistic regression") + 
  theme(legend.position = "none",axis.text = element_text(size = 12),
        plot.title = element_text(size = 16,hjust = 0.5)) + scale_x_log10()

P1
```

```{r}
P1
ggsave("odds_ratios_pdc_cluster_color.png", dpi = 1000,height = 6, width = 7)
```

