---
title: "pdc tot smoothing and pca"
output: html_notebook
---

set up:

```{r}
setwd("/Users/macbookpro/Documents/Applied Statistics/Project/Materiale_appstat/Code/Functional/")
library(tidyverse)
library(fda)
library(fields)
```

```{r}
load("~/Documents/Applied Statistics/Project/Materiale_appstat/Code/Functional/pazienti.RData")
load("~/Documents/Applied Statistics/Project/Materiale_appstat/Code/Functional/PDC_TOT.RData")
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/data.RData")
```

```{r}
n=dim(PDC_TOT)[2]
index=sample(n,20)
matplot(PDC_TOT[,index],type = 'l',main='dayly PDC',xlab='day',ylab='PDC')
```


## smoothing

### splines:

```{r}
days <- 1:365
basis <- create.bspline.basis(rangeval=range(days),nbasis=150,norder = 4)
data.fd <- Data2fd(y = PDC_TOT,argvals = days,basisobj = basis)
plot.fd(data.fd[index],xlab='days',ylab='PDC')
```


```{r}
basis <- create.bspline.basis(rangeval=range(days),nbasis=365,norder = 4)
data.fd <- Data2fd(y = PDC_TOT,argvals = days,basisobj = basis)
plot.fd(data.fd[index],xlab='days',ylab='PDC')
```

stability problem for large number of basis

```{r}
basis <- create.bspline.basis(rangeval=range(days),nbasis=200,norder = 4)
data.fd <- Data2fd(y = PDC_TOT,argvals = days,basisobj = basis)
plot.fd(data.fd[index],xlab='days',ylab='PDC')
```

choosing the best number of basis via gcv:

```{r}
n <- dim(PDC_TOT)[2]
nbasis <- seq(250,350,25)
gcv <- numeric(length(nbasis))
PDC_TOT_subset <- PDC_TOT[,sample(n,1000)]

for (i in 1:length(nbasis)){
  basis <- create.bspline.basis(rangeval=range(days),nbasis=nbasis[i],norder = 4)
  for(j in 1:1000){
    gcv[i] <- gcv[i] + smooth.basis(days, PDC_TOT_subset[,j], basis)$gcv
  }
  gcv[i] <- gcv[i]/n
}
```


```{r}
plot(nbasis,gcv,type = "b")
nbasis[which.min(gcv)]
abline(v = nbasis[which.min(gcv)], col = 2)
```

the best numver of basis is 250

fitting the data
```{r}
basis <- create.bspline.basis(rangeval=range(days),nbasis=250,norder = 4)
data.fd <- Data2fd(y = PDC_TOT,argvals = days,basisobj = basis)
```

since we are interested in doing the pcca let's look at the covatiance matrix 

```{r}
basis.1 <- create.bspline.basis(rangeval=range(days),nbasis=365,norder = 4)
data_W.fd.1 <- Data2fd(y = PDC_TOT,argvals = days,basisobj = basis.1)


basis.2 <- create.bspline.basis(rangeval=range(days),nbasis=250,norder = 4)
data_W.fd.2 <- Data2fd(y = PDC_TOT,argvals = days,basisobj = basis.2)

par(mfrow=c(1,2))

eval.1 <- eval.fd(days,data_W.fd.1)
image.plot(days,days,(cov(t(eval.1))[1:365,])) 

eval.2 <- eval.fd(days,data_W.fd.2)
image.plot(days,days,(cor(t(eval.2))[1:365,]))
```

the coariance sructure is very similar

fitting the pca on this:

```{r}
pca.res <- pca.fd(data.fd,nharm=10,centerfns=TRUE)
```

scree plot for the first 20 comp
```{r}
k <- 20
par(mfrow = c(1,2))
plot(pca.res$values[1:k ],xlab='number of components',ylab='Eigenvalues',type='b')
plot(cumsum(pca.res$values[1:k ])/sum(pca.res$values), type='b', axes=F, xlab='number of components', 
     ylab='contribution to the total variance', ylim=c(0.45,1))
abline(h=1, col='blue')
abline(h=0.9, lty=2, col='blue')
box()
axis(2,at=0:10/10,labels=0:10/10)
axis(1,at=1:k ,labels=1:k ,las=1)
mtext("scree plot f-PCA",outer = TRUE,side = 3,line = -1.5,cex = 1.5)
```

```{r}
cumsum(pca.res$values[1:10])/sum(pca.res$values)
```

as a perturbation of the mean:
```{r}
media <- mean.fd(data.fd)

par(mfrow = c(2,3))
plot(media,lwd=2,ylim=c(0,0.8),xlab='days',main='FPC1')
lines(media+pca.res$harmonics[1,]*sqrt(pca.res$values[1]), col=2)
lines(media-pca.res$harmonics[1,]*sqrt(pca.res$values[1]), col=3)

plot(media,lwd=2,ylim=c(0,0.75),xlab='days',main='FPC2')
lines(media+pca.res$harmonics[2,]*sqrt(pca.res$values[2]), col=2)
lines(media-pca.res$harmonics[2,]*sqrt(pca.res$values[2]), col=3)

plot(media,lwd=2,ylim=c(0,0.6),xlab='days',main='FPC3')
lines(media+pca.res$harmonics[3,]*sqrt(pca.res$values[3]), col=2)
lines(media-pca.res$harmonics[3,]*sqrt(pca.res$values[3]), col=3)

plot(media,lwd=2,ylim=c(0,0.6),xlab='days',main='FPC4')
lines(media+pca.res$harmonics[4,]*sqrt(pca.res$values[4]), col=2)
lines(media-pca.res$harmonics[4,]*sqrt(pca.res$values[4]), col=3)

plot(media,lwd=2,ylim=c(0.1,0.6),xlab='days',main='FPC5')
lines(media+pca.res$harmonics[5,]*sqrt(pca.res$values[5]), col=2)
lines(media-pca.res$harmonics[5,]*sqrt(pca.res$values[5]), col=3)

plot(media,lwd=2,ylim=c(0,0.6),xlab='days',main='FPC6')
lines(media+pca.res$harmonics[6,]*sqrt(pca.res$values[6]), col=2)
lines(media-pca.res$harmonics[6,]*sqrt(pca.res$values[6]), col=3)
```


```{r}
par(mfrow=c(2,3))
plot(pca.res, nx=50, pointplot=TRUE, harm=1:6, expand=0, cycle=FALSE,xlab='day')
```

we can clearly see that he components are different number of oscillations in he data
we coul use regularized pca if we wanr smoother pc but it seems we dont need them

looking at he scores:

```{r}
scores <- data.frame(pca.res$scores)
colnames(scores) <- c("comp1","comp2","comp3","comp4","comp5","comp6","comp7","comp8","comp9","comp10")
scores$label <- pazienti$labelOUT
```


```{r}
scores %>% ggplot(aes(comp1,comp2)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp1,comp3)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp2,comp3)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp1,comp4)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp1,comp5)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp2,comp4)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp3,comp4)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp2,comp5)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp3,comp5)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp4,comp5)) + geom_point(aes(col = label))
```

```{r}
scores %>% ggplot(aes(comp1,comp2)) + geom_point(aes(col = label)) + facet_wrap(vars(label))
```


```{r}
scores %>% ggplot(aes(y = comp1, fill = label)) + geom_histogram(bins = 30, position = "fill")
scores %>% ggplot(aes(y = comp2, fill = label)) + geom_histogram(bins = 30, position = "fill")
scores %>% ggplot(aes(y = comp3, fill = label)) + geom_histogram(bins = 30, position = "fill")
scores %>% ggplot(aes(y = comp4, fill = label)) + geom_histogram(bins = 30, position = "fill")
scores %>% ggplot(aes(y = comp5, fill = label)) + geom_histogram(bins = 30, position = "fill")
scores %>% ggplot(aes(y = comp6, fill = label)) + geom_histogram(bins = 30, position = "fill")
```

this are not easily separable

logistic regression on the pc components

```{r}
codes <- pazienti$CODREG
retained_patients <- data[data$COD_REG %in% codes,]
retained_patients$fpca1 <- pca.res$scores[,1]
retained_patients$fpca2 <- pca.res$scores[,2]
retained_patients$fpca3 <- pca.res$scores[,3]
```

```{r}
df <- retained_patients[,c(3,6,7,8,14,20,21,22)]
```

stratified k_fold cross-validation on labelout

```{r}
set.seed(123)
df_split <- initial_split(df, strata = labelOUT)
df_train <- training(df_split)
df_test <- testing(df_split)

set.seed(234)
df_folds <- vfold_cv(df_train, strata = labelOUT)

```

```{r}
glm_spec <- logistic_reg(engine = "glm")

rec_paper <- recipe(labelOUT ~ ., data = df_train) %>% 
  step_normalize(all_numeric_predictors()) %>% step_dummy(all_nominal_predictors())

wf_paper <- workflow(rec_paper,glm_spec)
```

```{r}
perf <- metric_set(accuracy,
                   sensitivity,
                   specificity,
                   precision,
                   recall,
                   roc_auc)

doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_basic <- fit_resamples(wf_paper, df_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_basic)
```

```{r}
collect_predictions(rs_basic) %>% group_by(id) %>% roc_curve(labelOUT,.pred_DECEDUTO) %>% autoplot()
```

fitting the model to the train data data

```{r}
fit_train <- fit(wf_paper,df_train)
```

looking at the coefficients:

```{r}
tidy(fit_train)
```

all of the coefficients are significant

looking at the odds ratios:

```{r}
tidy(fit_train,exponentiate=TRUE)
```

we can see that the ODDS ratio relative to the PDC_TOT is 1.94

### evaluation on test data

```{r}
test_fit <- last_fit(wf_paper,df_split,metrics = perf)
test_fit %>% collect_metrics()
```
the performance does not change o the test data, no overfitting.

let's investigate more:

```{r}
df_preds <- fit_train %>% augment(new_data = df_test)
```

```{r}
df_preds %>% roc_curve(labelOUT,.pred_DECEDUTO) %>% autoplot()
```

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```


smoothing splines:

```{r}
basis <- create.bspline.basis(days, norder=4)
lam <- 1e-5
```

```{r}
functionalPar <- fdPar(fdobj=basis, Lfdobj=2, lambda=lam)  
data.smooth <- smooth.basis(days, PDC_TOT, functionalPar)
eval.0 <- eval.fd(days, data.smooth$fd, Lfd=0)
```

```{r}
plot.fd(data.smooth$fd[index],xlab='days',ylab='PDC')
```

this looks the best, dooing gcv for lambad:

```{r}
lambda <- 10^seq(-12,-5,by = 0.5)
gcv <- numeric(length(lambda))
basis <- create.bspline.basis(days, norder=4)

for (i in 1:length(lambda)){
  functionalPar <- fdPar(fdobj=basis, Lfdobj=2, lambda=lambda[i])  
  data.smooth <- smooth.basis(days, PDC_TOT, functionalPar)
  gcv[i] <- mean(data.smooth$gcv)
}
```

```{r}
plot(log(lambda),gcv,type = "b")
abline(v = log(lambda[which.min(gcv)]), col = 2)
```

```{r}
lambda[which.min(gcv)]
```

the best lambda is s

```{r}
basis <- create.bspline.basis(days, norder=4)
lam <- 1e-8
functionalPar <- fdPar(fdobj=basis, Lfdobj=2, lambda=lam)  
data.smooth <- smooth.basis(days, PDC_TOT, functionalPar)
eval.0 <- eval.fd(days, data.smooth$fd, Lfd=0)
```

```{r}
plot.fd(data.smooth$fd[index],xlab='days',ylab='PDC')
```
this looks to be the best

```{r}
id <- sample(n,250)
plot.fd(data.smooth$fd[id],xlab='days',ylab='PDC',col = pazienti$labelOUT[id])
```

probably still difficult to classify


fitting the pca on this:

```{r}
pca.res <- pca.fd(data.smooth$fd,nharm=10,centerfns=TRUE)
```

scree plot for the first 20 comp
```{r}
k <- 20
par(mfrow = c(1,2))
plot(pca.res$values[1:k ],xlab='number of components',ylab='Eigenvalues',type='b')
plot(cumsum(pca.res$values[1:k ])/sum(pca.res$values), type='b', axes=F, xlab='number of components', 
     ylab='contribution to the total variance', ylim=c(0.45,1))
abline(h=1, col='blue')
abline(h=0.9, lty=2, col='blue')
box()
axis(2,at=0:10/10,labels=0:10/10)
axis(1,at=1:k ,labels=1:k ,las=1)
mtext("scree plot f-PCA",outer = TRUE,side = 3,line = -1.5,cex = 1.5)
```

```{r}
cumsum(pca.res$values[1:10])/sum(pca.res$values)
```

as a perturbation of the mean:
```{r}
media <- mean.fd(data.fd)

par(mfrow = c(2,3))
plot(media,lwd=2,ylim=c(0,0.9),xlab='days',main='FPC1')
lines(media+pca.res$harmonics[1,]*sqrt(pca.res$values[1]), col=2)
lines(media-pca.res$harmonics[1,]*sqrt(pca.res$values[1]), col=3)

plot(media,lwd=2,ylim=c(0,0.75),xlab='days',main='FPC2')
lines(media+pca.res$harmonics[2,]*sqrt(pca.res$values[2]), col=2)
lines(media-pca.res$harmonics[2,]*sqrt(pca.res$values[2]), col=3)

plot(media,lwd=2,ylim=c(0,0.6),xlab='days',main='FPC3')
lines(media+pca.res$harmonics[3,]*sqrt(pca.res$values[3]), col=2)
lines(media-pca.res$harmonics[3,]*sqrt(pca.res$values[3]), col=3)

plot(media,lwd=2,ylim=c(0,0.6),xlab='days',main='FPC4')
lines(media+pca.res$harmonics[4,]*sqrt(pca.res$values[4]), col=2)
lines(media-pca.res$harmonics[4,]*sqrt(pca.res$values[4]), col=3)

plot(media,lwd=2,ylim=c(0.1,0.6),xlab='days',main='FPC5')
lines(media+pca.res$harmonics[5,]*sqrt(pca.res$values[5]), col=2)
lines(media-pca.res$harmonics[5,]*sqrt(pca.res$values[5]), col=3)

plot(media,lwd=2,ylim=c(0,0.6),xlab='days',main='FPC6')
lines(media+pca.res$harmonics[6,]*sqrt(pca.res$values[6]), col=2)
lines(media-pca.res$harmonics[6,]*sqrt(pca.res$values[6]), col=3)
```

a better plot:
```{r}
mediaval <- eval.fd(days, media, Lfd=0)
mediapluspc1 <- eval.fd(days,media+pca.res$harmonics[1,]*sqrt(pca.res$values[1]),Lfd=0)
mediaminuspc1 <- eval.fd(days,media-pca.res$harmonics[1,]*sqrt(pca.res$values[1]),Lfd=0)

mediapluspc2 <- eval.fd(days,media+pca.res$harmonics[2,]*sqrt(pca.res$values[2]),Lfd=0)
mediaminuspc2 <- eval.fd(days,media-pca.res$harmonics[2,]*sqrt(pca.res$values[2]),Lfd=0)

mediapluspc3 <- eval.fd(days,media+pca.res$harmonics[3,]*sqrt(pca.res$values[3]),Lfd=0)
mediaminuspc3 <- eval.fd(days,media-pca.res$harmonics[3,]*sqrt(pca.res$values[3]),Lfd=0)
```

```{r}
nplots <- 9
#indexes_plot <- sample(n,nplots)

unicurves <- t(cbind(mediaval,mediapluspc1,mediaminuspc1,mediaval,mediapluspc2,mediaminuspc2,
                     mediaval,mediapluspc3,mediaminuspc3))

grids <- matrix(rep(1:365,nplots),nrow = nplots, ncol = 365,byrow = TRUE)
componen <- c(1,1,1,2,2,2,3,3,3)
col <- as.factor(c(1,2,3,1,2,3,1,2,3))
mycolors <- c("gray0","#F8766D" ,"#00B0F6")

wrangled_data <- tibble::tibble(
      grid = purrr::array_tree(grids, margin = 1),
      value = purrr::array_tree(unicurves, margin = 1),
      membership = componen,
      curve_id = 1:nplots,
      color = col
    ) |>
      tidyr::unnest(cols = c(grid, value)) |> 
  dplyr::bind_rows() |>
    dplyr::mutate(
      membership = as.factor(membership),
      curve_id = as.factor(curve_id),
    )

```

```{r}
P1 <- wrangled_data |>
      ggplot2::ggplot(ggplot2::aes(
        x = grid,
        y = value,
        color = color,
        group = curve_id
      )) +
      ggplot2::geom_line(linewidth = 0.9) +
      ggplot2::facet_grid(
        cols = ggplot2::vars(membership)
      ) +
      ggplot2::labs(
        x = "Day",
        y = "Running PDC"
      ) +
  scale_color_manual(values=mycolors) +
      ggplot2::theme(legend.position = "none",axis.title.y = element_blank(),
                     axis.title.x = element_text(size = 13))

P1
```

```{r}
P1
ggsave("fpca_comp.png", dpi = 1000,width = 10,height = 3)
```


```{r}
par(mfrow=c(2,3))
plot(pca.res, nx=50, pointplot=TRUE, harm=1:6, expand=0, cycle=FALSE,xlab='day')
```




we can clearly see that he components are different number of oscillations in he data
we coul use regularized pca if we wanr smoother pc but it seems we dont need them

looking at he scores:

```{r}
scores <- data.frame(pca.res$scores)
colnames(scores) <- c("comp1","comp2","comp3","comp4","comp5","comp6","comp7","comp8","comp9","comp10")
scores$label <- pazienti$labelOUT
```


```{r}
scores %>% ggplot(aes(comp1,comp2)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp1,comp3)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp2,comp3)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp1,comp4)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp1,comp5)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp2,comp4)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp3,comp4)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp2,comp5)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp3,comp5)) + geom_point(aes(col = label))
scores %>% ggplot(aes(comp4,comp5)) + geom_point(aes(col = label))
```

```{r}
scores %>% ggplot(aes(comp1,comp2)) + geom_point(aes(col = label)) + facet_wrap(vars(label))
```


```{r}
scores %>% ggplot(aes(y = comp1, fill = label)) + geom_histogram(bins = 30, position = "fill")
scores %>% ggplot(aes(y = comp2, fill = label)) + geom_histogram(bins = 30, position = "fill")
scores %>% ggplot(aes(y = comp3, fill = label)) + geom_histogram(bins = 30, position = "fill")
scores %>% ggplot(aes(y = comp4, fill = label)) + geom_histogram(bins = 30, position = "fill")
scores %>% ggplot(aes(y = comp5, fill = label)) + geom_histogram(bins = 30, position = "fill")
scores %>% ggplot(aes(y = comp6, fill = label)) + geom_histogram(bins = 30, position = "fill")
```

this are not easily separable

