---
title: "Exploratory data analysis"
output: html_notebook
---

this is a fisrt exploratory data analysis of the preprocessed data

```{r}
library(tidyverse)
library(GGally)
```

loading the preprocessed dataframes
```{r }
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/final_comorbidita.RData")
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/final_data.RData")
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/final_hospitalization.RData")
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/final_PDC.RData")
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/final_prescriptions.RData")
```

```{r }
glimpse(final_data)
final_data[,9:28] <- lapply(final_data[,9:28] , factor)
```

```{r}
glimpse(final_hospitalization)
```

```{r}
glimpse(final_prescriptions)
```

lets start by potting the distribution of different variables

### labelOUT

```{r}
ggplot(final_data,aes(labelOUT)) + geom_bar()
```

```{r}
final_data$labelOUT <- droplevels(final_data$labelOUT)
table(final_data$labelOUT)
```

the datasets is quite unbalanced, we have 5 times more truncated than not deceased people
we neet to keep this in mind when training models

### data_rif_ev

```{r}
 ggplot(final_data,aes(data_rif_ev,colour = labelOUT)) + geom_freqpoly(binwidth = 7)
```

we can see that the patients do not enter the study uniformly
the number of deaths goes down ad the times goes on as expected since the patients have less time to die

###SESSO

```{r}
ggplot(final_data, aes(SESSO)) + geom_bar(aes(fill = labelOUT))
```

the data seems quite balanced in terms of both the genders and the proportion of deaths per gender

```{r}
table(final_data$SESSO,final_data$labelOUT)
```

the proportion is very similar

### ASL_RESIDENZA

```{r}
ggplot(final_data, aes(ASL_RESIDENZA)) + geom_bar(aes(fill = labelOUT))
```

the patients are not uniformly distributed in the ASL

```{r}
table(final_data$labelOUT,final_data$ASL_RESIDENZA)
```

```{r}
sc = rep(0,30)
for(i in 1:15){
  sc[2*i - 1] = sum(table(final_data$labelOUT,final_data$ASL_RESIDENZA)[,i])
  sc[2*i] = sum(table(final_data$labelOUT,final_data$ASL_RESIDENZA)[,i])
}
table(final_data$labelOUT,final_data$ASL_RESIDENZA)/sc
```

the proportion of deeceeased is very closse for the different ASL

### data_studio_out

```{r}
final_data %>% filter(labelOUT == "DECEDUTO") %>%  ggplot(aes(data_studio_out)) + geom_freqpoly(binwidth = 7)
```

we can see that more patients die as the time goes on

### eta_min

```{r}
ggplot(final_data, aes(eta_Min)) + geom_histogram(aes(fill = labelOUT),binwidth = 1)
```
we can clearly see that the older the patients, the hihgher the proportion of deaths, we can also see that for very hold patientss we have a few data points

```{r}
sc = rep(0,50)
for(i in 1:50){
  sc[2*i - 1] = sum(table(final_data$labelOUT,final_data$eta_Min)[,i])
  sc[2*i] = sum(table(final_data$labelOUT,final_data$eta_Min)[,i])
}
age_table <- table(final_data$labelOUT,final_data$eta_Min)/sc
age_table
```

```{r}
y <- age_table[1,]
x = seq(50,99)
df = tibble(x,y)
ggplot(df,aes(x,y)) + geom_line()
```

### time_in_hosspital

```{r}
ggplot(final_data, aes(time_in_hospital,fill = labelOUT)) + geom_histogram(,binwidth = 1) + xlim(0,100)
```


```{r}
n = dim(table(final_data$labelOUT,final_data$time_in_hospital))[2]
sc = rep(0,n)
for(i in 1:n){
  sc[2*i - 1] = sum(table(final_data$labelOUT,final_data$time_in_hospital)[,i])
  sc[2*i] = sum(table(final_data$labelOUT,final_data$time_in_hospital)[,i])
}
h_table <- table(final_data$labelOUT,final_data$time_in_hospital)/sc
h_table
```


```{r}
y <- h_table[1,]
x = sort(unique(final_data$time_in_hospital))
df = tibble(x,y)
ggplot(df,aes(x,y)) + geom_line()
```

we can see that as the number of days in the hospital increases, the likelyhood of death increases

### comorbidità

```{r}
ggplot(final_data,aes(as.factor(comorbidita))) + geom_bar(aes(fill = labelOUT))
```

```{r}
n = dim(table(final_data$labelOUT,final_data$comorbidita))[2]
sc = rep(0,n)
for(i in 1:n){
  sc[2*i - 1] = sum(table(final_data$labelOUT,final_data$comorbidita)[,i])
  sc[2*i] = sum(table(final_data$labelOUT,final_data$comorbidita)[,i])
}
c_table <- table(final_data$labelOUT,final_data$comorbidita)/sc
c_table
```

```{r}
y <- c_table[1,]
x = sort(unique(final_data$comorbidita))
df = tibble(x,y)
ggplot(df,aes(x,y)) + geom_line()
```

we need to compute the carlston index and not just the number of comorbidities to have a better picture

### ICD

```{r}
ggplot(final_data,aes(as.factor(ICD))) + geom_bar(aes(fill = labelOUT))
```

```{r}
table(final_data$labelOUT,final_data$ICD)
```

### SHOCK

```{r}
ggplot(final_data,aes(as.factor(SHOCK))) + geom_bar(aes(fill = labelOUT))
```

```{r}
table(final_data$labelOUT,final_data$SHOCK)
```

### CABG

```{r}
ggplot(final_data,aes(as.factor(CABG))) + geom_bar(aes(fill = labelOUT))
```

```{r}
table(final_data$labelOUT,final_data$CABG)
```

### CABG

```{r}
ggplot(final_data,aes(as.factor(CABG))) + geom_bar(aes(fill = labelOUT))
```

```{r}
table(final_data$labelOUT,final_data$CABG)
```

### PTCA

```{r}
ggplot(final_data,aes(as.factor(PTCA))) + geom_bar(aes(fill = labelOUT))
```

```{r}
table(final_data$labelOUT,final_data$PTCA)
```

## adherence measures

### PDC_AA

```{r}
 ggplot(final_data,aes(PDC_AA,fill = labelOUT)) + geom_boxplot()
```

```{r}
 ggplot(final_data,aes(labelOUT,PDC_AA)) + geom_violin(aes(fill = labelOUT))
```

```{r}
 ggplot(final_data,aes(PDC_AA,fill = labelOUT)) + geom_histogram(binwidth = 0.05)
```


this does not look good

### PDC_BB

```{r}
 ggplot(final_data,aes(PDC_BB,fill = labelOUT)) + geom_boxplot()
```

```{r}
 ggplot(final_data,aes(labelOUT,PDC_BB)) + geom_violin(aes(fill = labelOUT))
```

```{r}
ggplot(final_data,aes(PDC_BB,fill = labelOUT)) + geom_histogram(binwidth = 0.05)
```

### PDC_RAS

```{r}
 ggplot(final_data,aes(PDC_RAS,fill = labelOUT)) + geom_boxplot()
```

```{r}
 ggplot(final_data,aes(labelOUT,PDC_RAS)) + geom_violin(aes(fill = labelOUT))
```

```{r}
ggplot(final_data,aes(PDC_RAS,fill = labelOUT)) + geom_histogram(binwidth = 0.05)
```


this looks promising

### n_drugs

```{r}
ggplot(final_data,aes(n_drugs)) + geom_bar(aes(fill= labelOUT))
```

```{r}
n = dim(table(final_data$labelOUT,final_data$n_drugs))[2]
sc = rep(0,n)
for(i in 1:n){
  sc[2*i - 1] = sum(table(final_data$labelOUT,final_data$n_drugs)[,i])
  sc[2*i] = sum(table(final_data$labelOUT,final_data$n_drugs)[,i])
}
z_table <- table(final_data$labelOUT,final_data$n_drugs)/sc
z_table
```
