---
title: "Survival_applied"
output: html_document
date: "2023-05-08"
---

```{r}
library(ISLR2)
load("~/Desktop/PoliMi 2022-2023/Secondo Semestre/Applied Statistics/Project/Survival/data_AA.RData")
load("~/Desktop/PoliMi 2022-2023/Secondo Semestre/Applied Statistics/Project/Survival/data_BB.RData")
load("~/Desktop/PoliMi 2022-2023/Secondo Semestre/Applied Statistics/Project/Survival/data_RAS.RData")
load("~/Desktop/PoliMi 2022-2023/Secondo Semestre/Applied Statistics/Project/Survival/data_AA_BB.RData")
load("~/Desktop/PoliMi 2022-2023/Secondo Semestre/Applied Statistics/Project/Survival/data_AA_RAS.RData")
load("~/Desktop/PoliMi 2022-2023/Secondo Semestre/Applied Statistics/Project/Survival/data_RAS_BB.RData")

#attach(data)
#table(sex)
#table(diagnosis) --> comorbidity in our case
#table(status) --> TRONCATO/DECEDUTO in labelOUT
```
# Survival data_RAS
```{r}
#attach(data_RAS)
data = data_RAS
attach(data)
```


The variables status and days_of_followup must be created:
```{r}
data$status <- ifelse(data$labelOUT == "TRONCATO",1, 0)
data$days_of_followup <- as.numeric(data_RAS$data_studio_out - data_RAS$data_rif_ev)
data$PAI_RANGE <- cut(
  data$PAI,
  breaks = c(0, 0.01, 0.34, 0.51, 0.67, 1.01),
  labels = c("0", "0.33.", "0.5", "0.67", "1"),
  right = FALSE
)
```
To begin the analysis, we re-create  the Kaplan-Meier survival curve  shown in  Figure 11.2, using the `survfit()` function within the `R` `survival` library. Here `time` corresponds to $y_i$, the time to the $i$th event (either censoring or death).
```{r}
attach(data)
library(survival)
fit.surv <- survfit(Surv(days_of_followup, status) ~ 1)
plot(fit.surv, xlab = "Days",
    ylab = "Estimated Probability of Survival", xlim = c(365,1095))

```

Next we create Kaplan-Meier survival curves that are stratified by `SESSO`.
```{r}
fit.sex <- survfit(Surv(days_of_followup, status) ~ SESSO)
plot.new()
plot(fit.sex, xlab = "Days",
    ylab = "Estimated Probability of Survival", col = c(2,4),xlim = c(365,1095))
#legend("bottomleft", levels(SESSO), col = c(2,4), lty = 1)
```
We can perform a log-rank test to compare the survival of males to females, using the `survdiff()` function.
```{r}
logrank.test <- survdiff(Surv(days_of_followup, status) ~ SESSO)
logrank.test
```
The resulting $p$-value is $0.8$, indicating  no evidence of a difference in survival between the two sexes.

Next, we fit  Cox proportional hazards models using the `coxph()`  function.
To begin, we consider a model that uses `SESSO` as the only predictor.
```{r}
fit.cox <- coxph(Surv(days_of_followup, status) ~ SESSO)
summary(fit.cox)
```
Regardless of which test we use, we see that there is no clear evidence for a difference in survival between males and females.

Now we fit a  model that makes use of additional predictors. 
```{r}
fit.all <- coxph(
Surv(days_of_followup, status) ~ SESSO + MCS + eta_Min + time_in_hospital + PAI_RANGE)
fit.all
```
Finally, we plot survival curves for each diagnosis category,  adjusting for the other predictors.
To make these plots, we set the values of the other predictors equal
to the mean for quantitative variables, and the modal value for
factors. We first create a data frame with four rows, one for each
level of diagnosis. The `survfit()` function will produce a curve for each of the rows in this data frame,
and one call to `plot()` will display them all in the same plot.
```{r}
modaldata <- data.frame(
     PAI_RANGE = levels(PAI_RANGE),
     SESSO = rep("F", 5),
     MCS = levels(MCS),
     time_in_hospital = rep(mean(time_in_hospital), 5),
     eta_Min = rep(mean(eta_Min), 5)
     )
survplots <- survfit(fit.all, newdata = modaldata)
{plot(survplots, xlab = "Days",ylab = "Survival Probability", col = 1:5, xlim = c(365,730))
legend("bottomleft", legend = levels(PAI_RANGE), col = 1:5, lty = 1)}

survplots <- survfit(fit.all, newdata = modaldata)
{plot(survplots, xlab = "Days",ylab = "Survival Probability", col = 1:5, xlim = c(365,1095), xaxt = "n")
legend("bottomleft", legend = levels(PAI_RANGE), col = 1:5, lty = 1)
axis(side = 1, at = c(365, 730, 1095), labels = c("1 Year", "2 Years","3 Years" ))}
```


