---
title: "drug comparison"
output: html_notebook
---

```{r}
library(tidyverse)
library(tidymodels)
library(themis)
library(patchwork)
```

## Explore the data

loading the data

```{r}
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/data.RData")
```

we want to build a logistic regression model using sesso, età, MCS, PDC, n_drugs

we need some preprocesing on this covariates

we filter the observation with only one drug to see the difference betwee the single drugs

```{r}
df <- data %>% filter(n_drugs == 1) %>%  select(c(SESSO,hospitalized,eta_Min,MCS,PDC_AA,PDC_BB,PDC_RAS))
```

we can now see the difference between drugs

```{r}
set.seed(123)
df_split <- initial_split(df, strata = hospitalized)
df_train <- training(df_split)
df_test <- testing(df_split)

set.seed(234)
df_folds <- vfold_cv(df_train, strata = hospitalized)
```

lets start by leaving the coefficients as numeric

```{r}
glm_spec <- logistic_reg(engine = "glm")

rec_paper <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min)

wf_paper <- workflow(rec_paper %>% step_downsample(hospitalized),glm_spec)
```

```{r}
perf <- metric_set(accuracy,
                   sensitivity,
                   specificity,
                   precision,
                   recall,
                   roc_auc)

doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_paper <- fit_resamples(wf_paper, df_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_paper)
```

```{r}
collect_predictions(rs_paper) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOPITALIZED) %>% autoplot()
```

on the test data

```{r}
test_fit <- wf_paper %>% last_fit(df_split,metrics = perf) 
```

```{r}
collect_metrics(test_fit)
```

```{r}
test_pred <- collect_predictions(test_fit)
```

```{r}
test_pred %>% conf_mat(hospitalized,.pred_class)
```

looking at the summary of the model

```{r}
summary(extract_fit_engine(test_fit))
```

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```


we can see that PDC_AA is not statisticly significant, in the case that we are taking only AA

let's try to reduce the mcs indexes to see it we get more accurate estimates of the coefficients

```{r}
df2 <- df
df2$MCS <- as.numeric(df2$MCS)
```

```{r}
set.seed(123)
df2_split <- initial_split(df2, strata = hospitalized)
df2_train <- training(df2_split)
df2_test <- testing(df2_split)

set.seed(234)
df2_folds <- vfold_cv(df2_train, strata = hospitalized)
```

```{r}
rec_nogender <- recipe(hospitalized ~ ., data = df2_train) %>% 
  step_normalize(eta_Min) %>% step_cut(MCS,breaks = c(1.5),include_outside_range = TRUE) 

wf_nogender<- workflow(rec_nogender %>% step_downsample(hospitalized),glm_spec)
```

```{r}
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_no_gendder <- fit_resamples(wf_nogender, df2_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_no_gendder)
```

```{r}
collect_predictions(rs_no_gendder) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

on the test data

```{r}
test_fit <- wf_nogender %>% last_fit(df2_split,metrics = perf) 
```

```{r}
collect_metrics(test_fit)
```

```{r}
test_pred <- collect_predictions(test_fit)
```

```{r}
test_pred %>% conf_mat(hospitalized,.pred_class)
```

looking at the summary of the model

```{r}
summary(extract_fit_engine(test_fit))
```

all of the predictors seem to be influential here

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```

this is confirmed, PDC_AA is not statistically significant

let's see if we consider the cat variables

```{r}
rec_cat <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% step_cut(starts_with("PDC"),breaks = c(0.75),include_outside_range = FALSE)

wf_cat <- workflow(rec_paper %>% step_downsample(hospitalized),glm_spec)
```

```{r}
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_cat <- fit_resamples(wf_cat, df_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_cat)
```

```{r}
collect_predictions(rs_cat) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

on the test data

```{r}
test_fit <- wf_cat %>% last_fit(df_split,metrics = perf) 
```

```{r}
collect_metrics(test_fit)
```

```{r}
test_pred <- collect_predictions(test_fit)
```

```{r}
test_pred %>% conf_mat(hospitalized,.pred_class)
```

looking at the summary of the model

```{r}
summary(extract_fit_engine(test_fit))
```

all of the predictors seem to be influential here

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```

the results do not change, AA does not seem statistically significant, even though it is close, we can see that they
all have a positive effect, if we consider only high adherence we ca see that all are effective

