---
title: "logistic regression my features"
output: html_notebook
---

```{r}
library(tidyverse)
library(tidymodels)
library(themis)
library(patchwork)
library(car)
```

```{r}
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/data.RData")
```

```{r}
df <- data %>% select(c(SESSO,hospitalized,eta_Min,MCS,PDC_TOT,n_drugs))
```

we consider n_drugs as a factor

```{r}
df$n_drugs <- as.factor(df$n_drugs)
df$MCS <- as.numeric(df$MCS)
```

stratified k_fold cross-validation on hospitalized

```{r}
set.seed(123)
df_split <- initial_split(df, strata = hospitalized)
df_train <- training(df_split)
df_test <- testing(df_split)

set.seed(234)
df_folds <- vfold_cv(df_train, strata = hospitalized)
```

we use different set of features in the processing of the data:
the MCS we use it at 1.5 to start with
the PDC_TOT at 0.5

age left as numeric for now

we use downsampling on the data for class imbalance

```{r}
glm_spec <- logistic_reg(engine = "glm")

rec_paper <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% step_cut(PDC_TOT,breaks = c(0.5),include_outside_range = FALSE) %>% 
  step_cut(MCS,breaks = c(1.5),include_outside_range = TRUE) 

wf_paper <- workflow(rec_paper %>% step_downsample(hospitalized),glm_spec)
```

```{r}
perf <- metric_set(accuracy,
                   sensitivity,
                   specificity,
                   precision,
                   recall,
                   roc_auc)

doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_paper <- fit_resamples(wf_paper, df_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_paper)
```

```{r}
collect_predictions(rs_paper) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

on the test data

```{r}
test_fit <- wf_paper %>% last_fit(df_split,metrics = perf) 
```

```{r}
collect_metrics(test_fit)
```

```{r}
test_pred <- collect_predictions(test_fit)
```

```{r}
test_pred %>% conf_mat(hospitalized,.pred_class)
```

looking at the summary of the model

```{r}
summary(extract_fit_engine(test_fit))
```

all of the predictors seem to be influential here

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```

another possibility is to let the MCS and the PDC be numeric and no segmentiation on the features

```{r}
rec_cat <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% step_cut(PDC_TOT,breaks = c(0.5),include_outside_range = FALSE) %>% 
  step_cut(MCS,breaks = c(1.5),include_outside_range = TRUE) 

rec_MCS_num <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% step_cut(PDC_TOT,breaks = c(0.5),include_outside_range = FALSE)

rec_PDC_num <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% step_cut(MCS,breaks = c(1.5),include_outside_range = TRUE) 

rec_both_num <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min)
```

```{r}
wf_set <-
  workflow_set(
    list(cat = rec_cat %>% step_downsample(hospitalized),
         MCS = rec_MCS_num %>% step_downsample(hospitalized),
         PDC = rec_PDC_num %>% step_downsample(hospitalized),
         both = rec_both_num %>% step_downsample(hospitalized)),
    list(glm = glm_spec)
  )
```

```{r}
doParallel::registerDoParallel()

res_cv <- wf_set %>% workflow_map("fit_resamples",resamples = df_folds,verbose = TRUE,
                                      metrics = perf, control = control_grid(save_workflow = TRUE, 
                                                                             save_pred = TRUE))
```

```{r}
autoplot(res_cv)
```

```{r}
collect_predictions(res_cv) %>% group_by(wflow_id) %>% roc_curve(hospitalized, .pred_NOT_HOSPITALIZED) %>% autoplot()
```

the difference is minimal between mcs left as number and the both of them left as munbers

```{r}
collect_metrics(res_cv) %>% filter(.metric == "roc_auc")
```


let's try to use finer detail in the cutting

```{r}
rec_simple <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% step_cut(PDC_TOT,breaks = c(0.4),include_outside_range = FALSE) %>% 
  step_cut(MCS,breaks = c(1.5),include_outside_range = TRUE)

rec_more_PDC <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% step_cut(PDC_TOT,breaks = c(0.25,0.75),include_outside_range = FALSE) %>% 
  step_cut(MCS,breaks = c(1.5),include_outside_range = TRUE)

rec_more_MCS <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% step_cut(PDC_TOT,breaks = c(0.4),include_outside_range = FALSE) %>% 
  step_cut(MCS,breaks = c(0.5,1.5,2.5,3.5),include_outside_range = TRUE)

rec_more_both <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% step_cut(PDC_TOT,breaks = c(0.25,0.75),include_outside_range = FALSE) %>% 
  step_cut(MCS,breaks = c(0.5,1.5,2.5,3.5),include_outside_range = TRUE)
```

```{r}
wf_set <-
  workflow_set(
    list(simple = rec_simple %>% step_downsample(hospitalized),
         PDC = rec_more_PDC %>% step_downsample(hospitalized),
         MS = rec_more_MCS %>% step_downsample(hospitalized),
         both = rec_more_both %>% step_downsample(hospitalized)),
    list(glm = glm_spec)
  )
```

```{r}
doParallel::registerDoParallel()

res_cv <- wf_set %>% workflow_map("fit_resamples",resamples = df_folds,verbose = TRUE,
                                      metrics = perf, control = control_grid(save_workflow = TRUE, save_pred = TRUE))
```

```{r}
autoplot(res_cv)
```

```{r}
collect_predictions(res_cv) %>% group_by(wflow_id) %>% roc_curve(hospitalized, .pred_NOT_HOSPITALIZED) %>% autoplot()
```

the bet model is the oe with segmented CCI, but more info in the PDC does not seem to make a difference

```{r}
collect_metrics(res_cv) %>% filter(.metric == "roc_auc")
```

it seem that the better prediction comes when having more info on the MCS, the PDC is not that relevant in this case

```{r}
summary(extract_fit_engine(fit(pull_workflow(res_cv,id = "MS_glm"),data = df_train)))
```

```{r}
summary(extract_fit_engine(fit(pull_workflow(res_cv,id = "both_glm"),data = df_train)))
```

in both cases it seems that the high pdc is positively linked to the hospitalization of the patient

we can see that all of the covariates seem significant

