---
title: "logistic regression PAI"
output: html_notebook
---

let's start by fitting a logistic regression on our data
starting from the total adherence measured via PAI

```{r}
library(tidyverse)
library(tidymodels)
library(themis)
library(patchwork)
```

loading the data

```{r}
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/data.RData")
```

```{r}
glimpse(data)
```

we want to build a logistic regression model using sesso, età, MCS, PAI, n_drugs

we need some preprocesing on this covariates

```{r}
df <- data %>% select(c(SESSO,hospitalized,eta_Min,MCS,PAI,n_drugs))
```

let's look at the PAI 

```{r}
ggplot(df,aes(as.factor(PAI))) + geom_bar(aes(fill = hospitalized))
```
I will keep it as a number, but possibly group it and use the groupped variable

first we keep this as a double, then we will group it

# logisic regression 

## the basic model

stratified k_fold cross-validation on hospitalized

```{r}
set.seed(123)
df_split <- initial_split(df, strata = hospitalized)
df_train <- training(df_split)
df_test <- testing(df_split)

set.seed(234)
df_folds <- vfold_cv(df_train, strata = hospitalized)

```

```{r}
glm_spec <- logistic_reg(engine = "glm")

rec_paper <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% 
  step_dummy(all_nominal_predictors())

wf_paper <- workflow(rec_paper %>% step_upsample(hospitalized),glm_spec)
```

```{r}
perf <- metric_set(accuracy,
                   sensitivity,
                   specificity,
                   precision,
                   recall,
                   roc_auc)

doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_basic <- fit_resamples(wf_paper, df_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_basic)
```
this is quite good: accucracy of 84% and ROC .77

```{r}
collect_predictions(rs_basic) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

fitting the model to the train data data

```{r}
fit_train <- fit(wf_paper,df_train)
```

looking at the coefficients:

```{r}
tidy(fit_train)
```

all of the coefficients are significant

looking at the odds ratios:

```{r}
tidy(fit_train,exponentiate=TRUE)
```

we can see that the ODDS ratio relative to the PAI is 0.65, which is quite high, keeping in mind that this is numeric here still


```{r}
test_fit <- last_fit(wf_paper,df_split,metrics = perf)
test_fit %>% collect_metrics()
```
the performance does not change o the test data, no overfitting.

let's investigate more:

```{r}
df_preds <- fit_train %>% augment(new_data = df_test)
```

```{r}
df_preds %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```
 
PAI is clearly significant
let's treat it as a factor and see if we can find differences

```{r}
df2 <- data %>% select(c(SESSO,hospitalized,eta_Min,MCS,PAI,n_drugs))
df2$PAI <- as.factor(df2$PAI)
```
 
```{r}
set.seed(123)
df2_split <- initial_split(df2, strata = hospitalized)
df2_train <- training(df2_split)
df2_test <- testing(df2_split)

set.seed(234)
df2_folds <- vfold_cv(df2_train, strata = hospitalized)
```
 
```{r}
rec2 <- recipe(hospitalized ~ ., data = df2_train) %>% 
  step_normalize(eta_Min) %>%
  step_dummy(all_nominal_predictors())

wf2 <- workflow(rec2 %>% step_upsample(hospitalized),glm_spec)
```

```{r}
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs2 <- fit_resamples(wf2, df2_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs2)
```
this is quite good: accucracy of 66% and ROC .74

```{r}
collect_predictions(rs2) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

fitting the model to the train data data

```{r}
fit_train <- fit(wf2,df2_train)
```

looking at the coefficients:

```{r}
tidy(fit_train)
```

all of the coefficients are significant

looking at the odds ratios:

```{r}
tidy(fit_train,exponentiate=TRUE)
```

we can see that the ODDS ratio relative to the PAI are low

### evaluation on test data

```{r}
test_fit <- last_fit(wf2,df2_split,metrics = perf)
test_fit %>% collect_metrics()
```
the performance does not change o the test data, no overfitting.

let's investigate more:

```{r}
df_preds <- fit_train %>% augment(new_data = df2_test)
```

```{r}
df_preds %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```

we can see that a high pai is clearly positively effecting the outcome of the patients

as far as 0.66 we have a high unceirtainty snce we have a few realizations with that value

let's gruop them into tree 0-0.33 0.5 0.66-1

```{r}
rec_3 <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% step_cut(PAI,breaks = c(0.4,0.6),include_outside_range = FALSE) %>% 
  step_dummy(all_nominal_predictors())

wf_3 <- workflow(rec_3 %>% step_upsample(hospitalized),glm_spec)
```

```{r}
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_3 <- fit_resamples(wf_3, df_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_3)
```
this is quite good: accucracy of 84% and ROC .77

```{r}
collect_predictions(rs_3) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

fitting the model to the train data data

```{r}
fit_train <- fit(wf_3,df_train)
```

looking at the coefficients:

```{r}
tidy(fit_train)
```

all of the coefficients are significant

looking at the odds ratios:

```{r}
tidy(fit_train,exponentiate=TRUE)
```

we can see that the ODDS ratio relative to the PAI is 0.75


```{r}
test_fit <- last_fit(wf_3,df_split,metrics = perf)
test_fit %>% collect_metrics()
```
the performance does not change o the test data, no overfitting.

let's investigate more:

```{r}
df_preds <- fit_train %>% augment(new_data = df_test)
```

```{r}
df_preds %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```
 
 it looks like they have the same effects, let's try a different grouping
 
```{r}
rec_3 <- recipe(hospitalized ~ ., data = df_train) %>% 
  step_normalize(eta_Min) %>% step_cut(PAI,breaks = c(0.1,0.4),include_outside_range = FALSE) %>% 
  step_dummy(all_nominal_predictors())

wf_3 <- workflow(rec_3 %>% step_upsample(hospitalized),glm_spec)
```

```{r}
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_3 <- fit_resamples(wf_3, df_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_3)
```
this is quite good: accucracy of 84% and ROC .77

```{r}
collect_predictions(rs_3) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

fitting the model to the train data data

```{r}
fit_train <- fit(wf_3,df_train)
```

looking at the coefficients:

```{r}
tidy(fit_train)
```

all of the coefficients are significant

looking at the odds ratios:

```{r}
tidy(fit_train,exponentiate=TRUE)
```

we can see that the ODDS ratio relative to the PAI is 0.8 and 0.73


```{r}
test_fit <- last_fit(wf_3,df_split,metrics = perf)
test_fit %>% collect_metrics()
```
the performance does not change o the test data, no overfitting.

let's investigate more:

```{r}
df_preds <- fit_train %>% augment(new_data = df_test)
```

```{r}
df_preds %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```
 
 even a small PAI is quite effective according to this model
 
 