---
title: "one grug at a time"
output: html_notebook
---

```{r}
library(tidyverse)
library(tidymodels)
library(themis)
library(patchwork)
```

## Explore the data

loading the data

```{r}
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/data.RData")
```

we want to build a logistic regression model using sesso, età, MCS, PDC, n_drugs

we need some preprocesing on this covariates

we compare taking at the drug in question (with or without others) ageinst the not taking it
I don't know if thiss makes sense clinically

# on the whole cohort

##AA

```{r}
dfAA <- data %>% select(c(SESSO,hospitalized,eta_Min,MCS,PDC_AA))
```

we can now see the difference between drugs

```{r}
set.seed(123)
dfAA_split <- initial_split(dfAA, strata = hospitalized)
dfAA_train <- training(dfAA_split)
dfAA_test <- testing(dfAA_split)

set.seed(234)
dfAA_folds <- vfold_cv(dfAA_train, strata = hospitalized)
```

```{r}
glm_spec <- logistic_reg(engine = "glm")

rec_AA <- recipe(hospitalized ~ ., data = dfAA_train) %>% 
  step_normalize(eta_Min) %>% 
  step_cut(starts_with("PDC"),breaks = c(0.001,0.25,0.5,0.75),include_outside_range = FALSE)

wf_AA <- workflow(rec_AA %>% step_downsample(hospitalized),glm_spec)
```

```{r}
perf <- metric_set(accuracy,
                   sensitivity,
                   specificity,
                   precision,
                   recall,
                   roc_auc)

doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_AA <- fit_resamples(wf_AA, dfAA_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_AA)
```

```{r}
collect_predictions(rs_AA) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

on the test data

```{r}
test_fit <- wf_AA %>% last_fit(dfAA_split,metrics = perf) 
```

```{r}
collect_metrics(test_fit)
```

```{r}
test_pred <- collect_predictions(test_fit)
```

```{r}
test_pred %>% conf_mat(hospitalized,.pred_class)
```

looking at the summary of the model

```{r}
summary(extract_fit_engine(test_fit))
```

all of the predictors seem to be influential here

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```

it looks like for low PDC it has a negative effect on the outcome, while in the high PDC it does not seem conclusive

we do the same for the other two drugs

##BB

```{r}
dfBB <- data %>% select(c(SESSO,hospitalized,eta_Min,MCS,PDC_BB))
```

we can now see the ifference between drugs

```{r}
set.seed(123)
dfBB_split <- initial_split(dfBB, strata = hospitalized)
dfBB_train <- training(dfBB_split)
dfBB_test <- testing(dfBB_split)

set.seed(234)
dfBB_folds <- vfold_cv(dfBB_train, strata = hospitalized)
```

lets start by leaving the coefficients as numeric

```{r}
rec_BB <- recipe(hospitalized ~ ., data = dfBB_train) %>% 
  step_normalize(eta_Min) %>% 
  step_cut(starts_with("PDC"),breaks = c(0.001,0.25,0.5,0.75),include_outside_range = FALSE)

wf_BB <- workflow(rec_BB %>% step_downsample(hospitalized),glm_spec)
```

```{r}
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_BB <- fit_resamples(wf_BB, dfBB_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_BB)
```

```{r}
collect_predictions(rs_BB) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

on the test data

```{r}
test_fit <- wf_BB %>% last_fit(dfBB_split,metrics = perf) 
```

```{r}
collect_metrics(test_fit)
```

```{r}
test_pred <- collect_predictions(test_fit)
```

```{r}
test_pred %>% conf_mat(hospitalized,.pred_class)
```

looking at the summary of the model

```{r}
summary(extract_fit_engine(test_fit))
```

all of the predictors seem to be influential here

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```

here we can see that for low adherance this is having a negative effect, while for high values it is decreasing the 
probability of hospitalization

##RAS

```{r}
dfRAS <- data %>% select(c(SESSO,hospitalized,eta_Min,MCS,PDC_RAS))
```

```{r}
set.seed(123)
dfRAS_split <- initial_split(dfRAS, strata = hospitalized)
dfRAS_train <- training(dfRAS_split)
dfRAS_test <- testing(dfRAS_split)

set.seed(234)
dfRAS_folds <- vfold_cv(dfRAS_train, strata = hospitalized)
```


```{r}
rec_RAS <- recipe(hospitalized ~ ., data = dfRAS_train) %>% 
  step_normalize(eta_Min) %>% 
  step_cut(starts_with("PDC"),breaks = c(0.001,0.25,0.5,0.75),include_outside_range = FALSE)

wf_RAS <- workflow(rec_RAS %>% step_downsample(hospitalized),glm_spec)
```

```{r}
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_RAS <- fit_resamples(wf_RAS, dfRAS_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_RAS)
```

```{r}
collect_predictions(rs_RAS) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

on the test data

```{r}
test_fit <- wf_RAS %>% last_fit(dfRAS_split,metrics = perf) 
```

```{r}
collect_metrics(test_fit)
```

```{r}
test_pred <- collect_predictions(test_fit)
```

```{r}
test_pred %>% conf_mat(hospitalized,.pred_class)
```

looking at the summary of the model

```{r}
summary(extract_fit_engine(test_fit))
```

all of the predictors seem to be influential here

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```

same as BB, here there seem not to be any positive effect at all

# looking only at the high adherance patients

to have a better comparison we look only at the patients that have high adherance to their respective protocols, ad see if the differences are the same also in this subset

##AA

```{r}
dfAA <- data %>% filter(PDC_TOT > 0.75) %>% select(c(SESSO,hospitalized,eta_Min,MCS,PDC_AA))
```

we can now see the difference between drugs

```{r}
set.seed(123)
dfAA_split <- initial_split(dfAA, strata = hospitalized)
dfAA_train <- training(dfAA_split)
dfAA_test <- testing(dfAA_split)

set.seed(234)
dfAA_folds <- vfold_cv(dfAA_train, strata = hospitalized)
```

```{r}
glm_spec <- logistic_reg(engine = "glm")

rec_AA <- recipe(hospitalized ~ ., data = dfAA_train) %>% 
  step_normalize(eta_Min) %>% 
  step_cut(starts_with("PDC"),breaks = c(0.75),include_outside_range = FALSE)

wf_AA <- workflow(rec_AA %>% step_downsample(hospitalized),glm_spec)
```

```{r}
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_AA <- fit_resamples(wf_AA, dfAA_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_AA)
```

```{r}
collect_predictions(rs_AA) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

on the test data

```{r}
test_fit <- wf_AA %>% last_fit(dfAA_split,metrics = perf) 
```

```{r}
collect_metrics(test_fit)
```

```{r}
test_pred <- collect_predictions(test_fit)
```

```{r}
test_pred %>% conf_mat(hospitalized,.pred_class)
```

looking at the summary of the model

```{r}
summary(extract_fit_engine(test_fit))
```

all of the predictors seem to be influential here

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```

it does not seem conclusive, but no evience in favour of incuding it 
we have a small number of datapoints that have high PDC_TOT and high pdc in this drug, this could create problem in the estimate

we do the same for the other two drugs

##BB

```{r}
dfBB <- data %>%  filter(PDC_TOT > 0.75) %>% select(c(SESSO,hospitalized,eta_Min,MCS,PDC_BB))
```

we can now see the ifference between drugs

```{r}
set.seed(123)
dfBB_split <- initial_split(dfBB, strata = hospitalized)
dfBB_train <- training(dfBB_split)
dfBB_test <- testing(dfBB_split)

set.seed(234)
dfBB_folds <- vfold_cv(dfBB_train, strata = hospitalized)
```

lets start by leaving the coefficients as numeric

```{r}
rec_BB <- recipe(hospitalized ~ ., data = dfBB_train) %>% 
  step_normalize(eta_Min) %>% 
  step_cut(starts_with("PDC"),breaks = c(0.75),include_outside_range = FALSE)

wf_BB <- workflow(rec_BB %>% step_downsample(hospitalized),glm_spec)
```

```{r}
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_BB <- fit_resamples(wf_BB, dfBB_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_BB)
```

```{r}
collect_predictions(rs_BB) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

on the test data

```{r}
test_fit <- wf_BB %>% last_fit(dfBB_split,metrics = perf) 
```

```{r}
collect_metrics(test_fit)
```

```{r}
test_pred <- collect_predictions(test_fit)
```

```{r}
test_pred %>% conf_mat(hospitalized,.pred_class)
```

looking at the summary of the model

```{r}
summary(extract_fit_engine(test_fit))
```

all of the predictors seem to be influential here

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```

also for BB we do not have enough evidence to say that it clearly improves the survival

id does not seem influential,but we have a few samples with this feature

##RAS

```{r}
dfRAS <- data %>% select(c(SESSO,hospitalized,eta_Min,MCS,PDC_RAS))
```

```{r}
set.seed(123)
dfRAS_split <- initial_split(dfRAS, strata = hospitalized)
dfRAS_train <- training(dfRAS_split)
dfRAS_test <- testing(dfRAS_split)

set.seed(234)
dfRAS_folds <- vfold_cv(dfRAS_train, strata = hospitalized)
```


```{r}
rec_RAS <- recipe(hospitalized ~ ., data = dfRAS_train) %>% 
  step_normalize(eta_Min) %>% 
  step_cut(starts_with("PDC"),breaks = c(0.75),include_outside_range = FALSE)

wf_RAS <- workflow(rec_RAS %>% step_downsample(hospitalized),glm_spec)
```

```{r}
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_RAS <- fit_resamples(wf_RAS, dfRAS_folds, control = ctrl_preds, metrics = perf)
```

```{r}
collect_metrics(rs_RAS)
```

```{r}
collect_predictions(rs_RAS) %>% group_by(id) %>% roc_curve(hospitalized,.pred_NOT_HOSPITALIZED) %>% autoplot()
```

on the test data

```{r}
test_fit <- wf_RAS %>% last_fit(dfRAS_split,metrics = perf) 
```

```{r}
collect_metrics(test_fit)
```

```{r}
test_pred <- collect_predictions(test_fit)
```

```{r}
test_pred %>% conf_mat(hospitalized,.pred_class)
```

looking at the summary of the model

```{r}
summary(extract_fit_engine(test_fit))
```

all of the predictors seem to be influential here

looking at the odds ratios:

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy(exponentiate = TRUE) %>% arrange(estimate) 
```

let's plot the coefficients

```{r}
test_fit %>% pull(.workflow) %>% pluck(1) %>% tidy() %>% 
  filter(term != "(Intercept)") %>%
  ggplot(aes(estimate, fct_reorder(term, estimate))) +
  geom_vline(xintercept = 0, color = "gray50", lty = 2, linewidth = 1.2) +
  geom_errorbar(aes(
    xmin = estimate - 1.96*std.error,
    xmax = estimate + 1.96*std.error),width = .7, alpha = 0.7) +
  geom_point(size = 2) +
  labs(y = NULL, x = "Coefficent from logistic regression")
```

here wee see that the clear difference starts when we have a PDC of more than 0.75

