---
title: "one feature at a time balanced data"
output: html_notebook
---

```{r}
library(tidyverse)
library(tidymodels)
library(themis)
library(parttree)
library(rpart.plot)
```

```{r}
load("~/Documents/Applied Statistics/Project/Materiale_appstat/final dataframes/data.RData")
```

focusing on the two features that we want to split into chunks

```{r}
df <- data %>% select(c(hospitalized,MCS,PDC_TOT,eta_Min))
```

balancing the data from the beginning

```{r}
df_balanced <- recipe(~., df) %>%
    step_downsample(hospitalized, under_ratio = 1) %>%
    prep() %>%
    bake(new_data = NULL)

table(df_balanced$hospitalized)
```

```{r}
set.seed(123)
df_split <- initial_split(df_balanced, strata = hospitalized)
df_train <- training(df_split)
df_test <- testing(df_split)

set.seed(234)
df_folds <- vfold_cv(df_train, strata = hospitalized)
```

we start only using PDC_TOT

```{r}
g1 <- ggplot(df, aes(PDC_TOT)) + geom_freqpoly(aes(colour = hospitalized),binwidth = 0.01) + 
  theme(legend.position = "bottom")
g2 <-df %>% ggplot(aes(y = PDC_TOT, fill = hospitalized)) + geom_histogram(binwidth = 0.01, position = "fill") + 
    theme(legend.position = "none")

g1 +g2
```

this probably will be difficult

```{r}
rec_basic <- recipe(hospitalized ~ PDC_TOT, data = df_train)
```

we want to split the PDC_TOT into at most 4 categories => tree_depth <= 3 (2 + root)

```{r}
tree_spec <- decision_tree(cost_complexity = 1e-3, tree_depth = 3, min_n = 500 ) %>% 
  set_mode("classification") %>% set_engine("rpart")
```

```{r}
perf <- metric_set(accuracy, sensitivity, specificity, roc_auc, precision, recall)
ctrl_preds <- control_resamples(save_pred = TRUE)
```

```{r}
test_rs <- last_fit(tree_spec,rec_basic,df_split,metrics = perf)
```

```{r}
collect_metrics(test_rs)
```

```{r}
rpart.plot(extract_fit_engine(test_rs),extra = 3,roundint = FALSE)
```


this feature seemed relevant also in the other experiment

let's try to build a deeper tree and see what happens

```{r}
tree_spec2 <- decision_tree(cost_complexity = 1e-3, tree_depth = 4, min_n = 400 ) %>% 
  set_mode("classification") %>% set_engine("rpart")
```

```{r}
test_rs <- last_fit(tree_spec2,rec_basic %>% step_downsample(hospitalized),df_split,metrics = perf)
```

```{r}
collect_metrics(test_rs)
```

```{r}
rpart.plot(extract_fit_engine(test_rs),extra = 3,roundint = FALSE)
```

probably we shoul pick 0.15 as the threshold

we do the same for MCS

```{r}
g1 <- ggplot(df,aes(MCS)) + geom_bar(aes(fill = hospitalized)) + theme(legend.position = "bottom")
g2 <- df %>% ggplot(aes(y = MCS, fill = hospitalized)) + geom_bar(position = "fill") + theme(legend.position = "none")

g1 + g2
```


```{r}
rec_basic <- recipe(hospitalized ~ MCS, data = df_train)
```

we want to split the MCS into at most 4 categories => tree_depth <= 3 (2 + root)

```{r}
tree_spec <- decision_tree(cost_complexity = 1e-1, tree_depth = 3, min_n = 500 ) %>% 
  set_mode("classification") %>% set_engine("rpart")
```

```{r}
fit_tree <- fit_resamples(tree_spec,rec_basic %>% step_downsample(hospitalized),df_folds,metrics = perf)
```

```{r}
collect_metrics(fit_tree)
```

```{r}
test_rs <- last_fit(tree_spec,rec_basic %>% step_downsample(hospitalized),df_split,metrics = perf)
```

```{r}
collect_metrics(test_rs)
```

```{r}
rpart.plot(extract_fit_engine(test_rs),extra = 3,roundint = FALSE)
```

it seems that the biggest step is for MCS = 0-1 or 2,3,4


let's try to build a deeper tree and see what happens

```{r}
tree_spec2 <- decision_tree(cost_complexity = 0.0, tree_depth = 5, min_n = 100) %>% 
  set_mode("classification") %>% set_engine("rpart")
```

```{r}
test_rs <- last_fit(tree_spec2,rec_basic ,df_split,metrics = perf)
```

```{r}
collect_metrics(test_rs)
```

```{r}
rpart.plot(extract_fit_engine(test_rs),extra = 3,roundint = FALSE)
```

there seem to be a difference between 0 1 and 2-3-4

now i want to create the a single decision boundary for the single drug usages to potentially use in the polytherapy interaction of the logistic regression

```{r}
df2 <- data %>% select(c(hospitalized,PDC_AA,PDC_BB,PDC_RAS)) 
df2_balanced <- recipe(~., df2) %>%
    step_downsample(hospitalized, under_ratio = 1) %>%
    prep() %>%
    bake(new_data = NULL)

table(df2_balanced$hospitalized)
```

```{r}
set.seed(123)
df2_split <- initial_split(df2_balanced, strata = hospitalized)
df2_train <- training(df2_split)
df2_test <- testing(df2_split)

set.seed(234)
df2_folds <- vfold_cv(df2_train, strata = hospitalized)
```

we start using PDC_AA

```{r}
g1 <- ggplot(df2, aes(PDC_AA)) + geom_histogram(aes(fill = hospitalized),binwidth = 0.05) + 
  theme(legend.position = "none")

g2 <- df2 %>% ggplot(aes(y = PDC_AA, fill = hospitalized)) + geom_histogram(binwidth = 0.01, position = "fill") +
  theme(legend.position = "none")

g1 + g2
```

this probably will be difficult

```{r}
rec_AA <- recipe(hospitalized ~ PDC_AA, data = df2_train)
```

we want to split the PDC_TOT into at most 4 categories => tree_depth <= 3 (2 + root)

```{r}
tree_spec <- decision_tree(cost_complexity = 1e-3, tree_depth = 3, min_n = 400 ) %>% 
  set_mode("classification") %>% set_engine("rpart")
```

```{r}
test_rs <- last_fit(tree_spec,rec_AA %>% step_downsample(hospitalized) ,df2_split,metrics = perf)
```

```{r}
collect_metrics(test_rs)
```

```{r}
rpart.plot(extract_fit_engine(test_rs),extra = 3,roundint = FALSE)
```

the threshols seems to be at 0.5

we start using PDC_BB

```{r}
g1 <- ggplot(df2, aes(PDC_BB)) + geom_histogram(aes(fill = hospitalized),binwidth = 0.05) + 
  theme(legend.position = "none")

g2 <- df2 %>% ggplot(aes(y = PDC_BB, fill = hospitalized)) + geom_histogram(binwidth = 0.01, position = "fill") +
  theme(legend.position = "none")

g1 + g2
```

this probably will be difficult

```{r}
rec_BB <- recipe(hospitalized ~ PDC_BB, data = df2_train)
```

we want to split the PDC_TOT into at most 4 categories => tree_depth <= 3 (2 + root)

```{r}
tree_spec <- decision_tree(cost_complexity = 0.0, tree_depth = 3, min_n = 500 ) %>% 
  set_mode("classification") %>% set_engine("rpart")
```

```{r}
test_rs <- last_fit(tree_spec,rec_BB %>% step_downsample(hospitalized) ,df2_split,metrics = perf)
```

```{r}
collect_metrics(test_rs)
```

```{r}
rpart.plot(extract_fit_engine(test_rs),extra = 3,roundint = FALSE)
```

the threshold seems 0.012, very low, the others are 0.52 and 0.86

we start using PDC_RAS

```{r}
g1 <- ggplot(df2, aes(PDC_RAS)) + geom_histogram(aes(fill = hospitalized),binwidth = 0.05) + 
  theme(legend.position = "none")

g2 <- df2 %>% ggplot(aes(y = PDC_RAS, fill = hospitalized)) + geom_histogram(binwidth = 0.01, position = "fill") +
  theme(legend.position = "none")

g1 + g2
```


```{r}
rec_RAS <- recipe(hospitalized ~ PDC_RAS, data = df2_train)
```

we want to split the PDC into at most 4 categories => tree_depth <= 3 (2 + root)

```{r}
tree_spec <- decision_tree(cost_complexity = 0.0, tree_depth = 3, min_n = 500 ) %>% 
  set_mode("classification") %>% set_engine("rpart")
```

```{r}
test_rs <- last_fit(tree_spec,rec_RAS %>% step_downsample(hospitalized) ,df2_split,metrics = perf)
```

```{r}
collect_metrics(test_rs)
```

```{r}
rpart.plot(extract_fit_engine(test_rs),extra = 3,roundint = FALSE)
```

here there seem to be a cutoff at either 82 or 0.94

for eta

```{r}
g1 <- ggplot(df, aes(eta_Min)) + geom_histogram(aes(fill = hospitalized),binwidth = 1) + 
  theme(legend.position = "none")

g2 <- df %>% ggplot(aes(y = eta_Min, fill = hospitalized)) + geom_histogram(binwidth = 1, position = "fill") +
  theme(legend.position = "none")

g1 + g2
```

```{r}
rec_eta <- recipe(hospitalized ~ eta_Min, data = df_train)
```

```{r}
tree_spec <- decision_tree(cost_complexity = 1e-3, tree_depth = 3, min_n = 100 ) %>% 
  set_mode("classification") %>% set_engine("rpart")
```

```{r}
test_rs <- last_fit(tree_spec,rec_eta %>% step_downsample(hospitalized) ,df_split,metrics = perf)
```

```{r}
collect_metrics(test_rs)
```

```{r}
rpart.plot(extract_fit_engine(test_rs),extra = 3,roundint = FALSE)
```

the step seems to be at 84

